---
title: "Geographical analysis of media"
subtitle: "6. Spatial Networks"
author: "Claude Grasland"
output: html_notebook
---


```{r setup4, echo = FALSE, comment = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = FALSE, warning = FALSE)
library(knitr)
library(dplyr)
library(data.table)
library(visNetwork, quietly = T)
library(sf)
```




# Networks


## Preparation of data

### Load Hypercube

```{r}
# Load complete hypercube
hc <- readRDS("data/corpus/hc_mycorpus_states_mobil_day.RDS")
hc <- hc %>% filter(where1 !="_no_", where2 != "_no_")

# Eliminate non foreign news (french news for french newspapers ...)
hc<-hc[hc$where1 != substr(who,4,6),]
hc<-hc[hc$where2 != substr(who,4,6),]

# Add complete labels
map<-readRDS("dict/world_map_4326.Rdata")
labs<-st_drop_geometry(map)
labs<-don[,c(1,4)]

# Shorten the name of USA
labs$NAME[labs$ISO3=="USA"]<-"U.S.A."


names(labs)<-c("where1","geofr1")
hc<-left_join(hc,labs)
names(labs)<-c("where2","geofr2")
hc<-left_join(hc,labs)

hc_geo_geo <- hc
```


### Hypercube Filter (function)

```{r}
hc_filter <- function(don = hc,
                      who = "who",
                      when = "when",
                      where1 = "where1",
                      where2 = "where2",
                      wgt = "tags",
                      self = FALSE,
                      when_start = NA,
                      when_end = NA,
                      who_exc = NA,
                      who_inc = NA,
                      where1_exc = NA,
                      where1_inc = NA,
                      where2_exc = NA,
                      where2_inc = NA)

  {                          
  
    df<-data.table(who = don[[who]],
                   when = don[[when]],
                   where1 = don[[where1]],
                   where2 = don[[where2]],
                   wgt = don[[wgt]])
    
    # Select time period
        if (is.na(when_start)==FALSE) { 
        df <- df[when >= as.Date(when_start), ]}
        if (is.na(when_end)==FALSE) { 
        df <- df[when <= as.Date(when_end), ]}
    # Select who
        if (is.na(who_exc)==FALSE) { 
        df <- df[!(who %in% who_exc), ]}
        if (is.na(who_inc)==FALSE) { 
        df <- df[(who %in% who_inc), ]}
    # Select where1
        if (is.na(where1_exc)==FALSE) { 
        df <- df[!(where1 %in% where1_exc), ]}
        if (is.na(where1_inc)==FALSE) { 
        df <- df[(where1 %in% where1_inc), ]}
    # Select where2
        if (is.na(where2_exc)==FALSE) { 
        df <- df[!(where2 %in% where2_exc), ]}
        if (is.na(where2_inc)==FALSE) { 
        df <- df[(where2 %in% where2_inc), ]}
    # eliminate internal links
       if (self==FALSE) { 
        df <- df[(where1 != where2), ]}
    return(df)
  
}
```

### Matrix builder (function)

```{r}
build_int <- function(don = don,       # a dataframe with columns i, j , Fij
                      i = "where1",
                      j = "where2",
                      Fij = "wgt",
                      s1 = 1,
                      s2 = 1,
                      n1 = 1,
                      n2 = 1,
                      k = 0)

{  
  df<-data.table(i=don[[i]],j=don[[j]],Fij=don[[Fij]])
  int <-df[,.(Fij=sum(Fij)),.(i,j)]
  int<-dcast(int,formula = i~j,fill = 0)
  mat<-as.matrix(int[,-1])
  row.names(mat)<-int$i
  mat<-mat[apply(mat,1,sum)>=s1,apply(mat,2,sum)>=s2 ]
  m0<-mat
  m0[m0<k]<-0
  m0[m0>=k]<-1
  mat<-mat[apply(m0,1,sum)>=n1,apply(m0,2,sum)>=n2 ]
  int<-reshape2::melt(mat)
  names(int) <-c("i","j","Fij")
  return(int)
}

```


### Random model (function)

```{r}

rand_int <- function(int = int, # A table with columns i, j Fij
                     maxsize = 100000,
                     diag    = FALSE,
                     resid   = FALSE) {
    # Eliminate diagonal ?
    if (diag==FALSE) { 
        int <- int[as.character(int$i) != as.character(int$j), ]}
  
    # Compute model if size not too large
    if (dim(int)[1] < maxsize) {
       # Proceed to poisson regression model
       mod <- glm( formula = Fij ~ i + j,family = "poisson", data = int)
  
       # Add residuals if requested
       if(resid == TRUE)   { 
          # Add estimates
          int$Eij <- mod$fitted.values

          # Add absolute residuals
          int$Rabs_ij <- int$Fij-int$Eij

          # Add relative residuals
          int$Rrel_ij <- int$Fij/int$Eij

          # Add chi-square residuals
          int$Rchi_ij <-  (int$Rabs_ij)**2 / int$Eij
          int$Rchi_ij[int$Rabs_ij<0]<- -int$Rchi_ij[int$Rabs_ij<0]
          }
         
    } else { paste ("Table > 100000 -  \n 
                     modify maxsize =  parameter \n
                     if you are sure that your computer can do it !")}
  # Export results
  int$i<-as.character(int$i)
  int$j<-as.character(int$j)
  return(int)
  
 }

```


### Visualize network (function)

This standard procedure has to be adapted to particular cases

```{r}
geo_network<- function(don = don,
                       from = "i",
                        to = "j", 
                        size = "Fij",
                        minsize = 1,
                        maxsize = NA,
                        test = "Fij",
                        mintest = 1,
                        loops  = FALSE, 
                        title = "Network")

{
int<-data.frame(i = as.character(don[,from]),
                j = as.character(don[,to]),
                size = don[,size],
                test = don[,test]
                )
if (is.na(minsize)==FALSE) {int =int[int$size >= minsize,]} 
if (is.na(maxsize)==FALSE) {int =int[int$size <= maxsize,]} 
if (is.na(mintest)==FALSE) {int =int[int$test >= mintest,]}

nodes<-data.frame(code = unique(c(int$i,int$j)))
nodes$code<-as.character(nodes$code)
nodes$id<-1:length(nodes$code)
nodes$label<-nodes$code
nodes$color <-"gray"
nodes$color[nodes$code %in% int$j]<-"red"


# Adjust edge codes
edges <- int %>% mutate(width = 5+5*size / max(size)) %>%
                left_join(nodes %>% select(i=code, from = id)) %>%  
                left_join(nodes %>% select(j=code, to = id )) 

# compute nodesize
toti<-int %>% group_by(i) %>% summarize(size =sum(size)) %>% select (code=i,size)
totj<-int %>% group_by(j) %>% summarize(size =sum(size)) %>% select (code=j,size)
tot<-rbind(toti,totj)
tot<-unique(tot)
tot$code<-as.factor(tot$code)
nodes <- left_join(nodes,tot) %>% mutate(value = 1 +5*sqrt(size/max(size)))


#sel_nodes <-nodes %>% filter(code %in% unique(c(sel_edges$i,sel_edges$j)))

# eliminate loops

if(loops == FALSE) {edges <- edges[edges$from < edges$to,]}

net<- visNetwork(nodes, 
                  edges, 
                  main = title,
height = "1000px", 
                  width = "70%")   %>%   
   visNodes(scaling =list(min =20, max=60, 
                          label=list(min=20,max=80, 
                                    maxVisible = 20)))%>%
  visEdges(scaling = list(min=20,max=60))%>%
       visOptions(highlightNearest = TRUE,
     #               selectedBy = "group", 
    #               manipulation = TRUE,
                  nodesIdSelection = TRUE) %>%
        visInteraction(navigationButtons = TRUE) %>%
         visLegend() %>%
      visIgraphLayout(layout ="layout.fruchterman.reingold",smooth = TRUE)

net
 return(net)
 } 

```


## Network analysis


### Reference network

We can firstly realize a network that combine all periods of time (when), all media (who) and all topics (what), which provide a reference for the next analysis.

```{r}
hc<-hc_geo_geo 



hc<-hc_filter(don = hc,
                             wgt = "tags",
                             where1 = "geofr1",
                             where2 = "geofr2",
                             where1_exc = c("_no_"),
                             where2_exc = c("_no_"),
                             self = FALSE
                           )

int <- build_int(don = hc,
                 s1=2,
                 s2=2,
                 n1=2,
                 n2=2,
                 k=0)

mod<-rand_int(int,
              resid = TRUE,
              diag = FALSE)


network<- geo_network(mod,
                      size = "Fij",
                      minsize = 30,
                      test = "Rchi_ij",
                      mintest = 3.84)
network

```
- **Comment** : The network is a combination of *structural proximity* (countries that are linked by relation not related to the period of observation) and *event linkage* related to specific events or crisis that has produced temporary connections generally not observed. The preferential linkage between Israël and Palestine is typically a structural linkage but the association between Liberia, Guinea and Sierra Leone has been induced by the Ebola outbreak. The strong linkage between Ukraine and Russia is a combination of the two types of effects. 



### Focus on a media

We can isolate a specific media like El Watan in order to examine to what extent his network is different fromthe reference network. But the parameters of selections has to be adapted because the size of the sample is smaller. 

```{r}
hc<-hc_geo_geo %>% filter(who == "fr_DZA_elwata")



hc<-hc_filter(don = hc,
                             wgt = "tags",
                             where1 = "geofr1",
                             where2 = "geofr2",
                             where1_exc = c("_no_"),
                             where2_exc = c("_no_"),
                             self = FALSE
                           )

int <- build_int(don = hc,
                 s1=2,
                 s2=2,
                 n1=1,
                 n2=1,
                 k=0)

mod<-rand_int(int,
              resid = TRUE,
              diag = FALSE)


network<- geo_network(mod,
                      size = "Fij",
                      minsize = 10,
                      test = "Rchi_ij",
                      mintest = 3.84)
network

```
- **Comments :** The network of prefernetial associations from El Watan is clearly more concentrated than the reference network on Mediterranean countries, Middle East and Africa. We notice specific relations related to geopolitical conflicts where Algeria is directly concerned (Morocco - Spain - Western Sahara). We can also notice that the so-called "European Migrant crisis" is less covered when it is related to Balkans.


### Focus on a topic

We can also propose a focus on the countries that has been specifically associated in relation with a topic. For example, the topic of migrants and refugees. We are oblige to reduce one more time the different thresholds as the number of news is very limited, but the result is interesting.

```{r}
hc<-hc_geo_geo %>% filter(what != "_no_")



hc<-hc_filter(don = hc,
                             wgt = "tags",
                             where1 = "geofr1",
                             where2 = "geofr2",
                             where1_exc = c("_no_"),
                             where2_exc = c("_no_"),
                             self = FALSE
                           )

int <- build_int(don = hc,
                 s1=1,
                 s2=1,
                 n1=1,
                 n2=1,
                 k=0)

mod<-rand_int(int,
              resid = TRUE,
              diag = FALSE)


network<- geo_network(mod,
                      size = "Fij",
                      minsize = 3,
                      test = "Rchi_ij",
                      mintest = 3.84)
network

```
- **Comments :** We can indeed recognize the different events related to the crossing of borders by migrants or refugees. We recognize without surprise the major events occuring in the Mediterranean. But the network reveals also othe crisis occuring in central America (miogrants from Cuba to USA) or in southern Asia (Rohyngias crisis)


